---
title: "Week 7 - Homework"
author: "STAT 420, Summer 2020, D. Unger"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---

***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.alin = "center")
```

## Exercise 1 (EPA Emissions Data)

For this exercise, we will use the data stored in [`epa2017.csv`](epa2017.csv). It contains detailed descriptions of vehicles manufactured in 2017 that were used for fuel economy testing [as performed by the Environment Protection Agency]( https://www.epa.gov/air-emissions-inventories). The variables in the dataset are:  

- `Make` - Manufacturer
- `Model` - Model of vehicle
- `ID` - Manufacturer defined vehicle identification number within EPA's computer system (not a VIN number)
- `disp` - Cubic inch displacement of test vehicle
- `type` - Car, truck, or both (for vehicles that meet specifications of both car and truck, like smaller SUVs or crossovers)
- `horse` - Rated horsepower, in foot-pounds per second
- `cyl` - Number of cylinders
- `lockup` - Vehicle has transmission lockup; N or Y
- `drive` - Drivetrain system code
    - A = All-wheel drive
    - F = Front-wheel drive
    - P = Part-time 4-wheel drive
    - R = Rear-wheel drive
    - 4 = 4-wheel drive
- `weight` - Test weight, in pounds
- `axleratio` - Axle ratio
- `nvratio` - n/v ratio (engine speed versus vehicle speed at 50 mph)
- `THC` - Total hydrocarbons, in grams per mile (g/mi)
- `CO` - Carbon monoxide (a regulated pollutant), in g/mi
- `CO2` - Carbon dioxide (the primary byproduct of all fossil fuel combustion), in g/mi
- `mpg` - Fuel economy, in miles per gallon

We will attempt to model `CO2` using both `horse` and `type`. In practice, we would use many more predictors, but limiting ourselves to these two, one numeric and one factor, will allow us to create a number of plots.

Load the data, and check its structure using `str()`. Verify that `type` is a factor; if not, coerce it to be a factor.
```{r}
epa2017 <- read.csv("epa2017.csv")

str(epa2017)
```


**(a)** Do the following:

- Make a scatterplot of `CO2` versus `horse`. Use a different color point for each vehicle `type`.
- Fit a simple linear regression model with `CO2` as the response and only `horse` as the predictor.
- Add the fitted regression line to the scatterplot. 

```{r fig.height=10, fig.width=10,echo=FALSE}
CO2_model_slr = lm(CO2 ~ horse, data = epa2017)

plot(CO2 ~ horse, data = epa2017, col = c(1,4,6), pch = c(1,4,6), cex = 2)
abline(CO2_model_slr, lwd = 3, col ="black")
legend("topright", c("Both", "Car", "Truck"), col = c(1,4,6), pch = c(1,4,6))

```

```{r, echo=FALSE}
summary(CO2_model_slr)$coef

predict(CO2_model_slr, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)
```
**Comment**

- The regression line for simple regression model seems to fit alright in the graph, showing a linear relationship betweenthe $CO2$ and $horsepower$. However this regression line doesn't show much of the relationship or impact of $type$ which I think it should be.

- The estimate for the average change in `CO2` for a one foot-pound per second increase in `horse` for a vehicle of type `car` is `r coef(CO2_model_slr)[2]`. 

- The interval for Subaru Impreza Wagon, which is a vehicle with 148 horsepower and is considered type `Both` in 90% prediction interval has lower bound as `r predict(CO2_model_slr, data.frame(horse = c(148)), interval = "prediction", level = 0.90)[2]` and upper bound as `r predict(CO2_model_slr, data.frame(horse = c(148)), interval = "prediction", level = 0.90)[3]`

**(b)** Do the following:

- Make a scatterplot of `CO2` versus `horse`. Use a different color point for each vehicle `type`.
- Fit an additive multiple regression model with `CO2` as the response and `horse` and `type` as the predictors.
- Add the fitted regression "lines" to the scatterplot with the same colors as their respective points (one line for each vehicle type)

```{r fig.height=10, fig.width=10, echo=FALSE}
CO2_model_add = lm(CO2 ~ horse + type, data = epa2017)

#intercept
int_both = coef(CO2_model_add)[1]
int_car = coef(CO2_model_add)[1] + coef(CO2_model_add)[3]
int_truck = coef(CO2_model_add)[1] + coef(CO2_model_add)[4]

#slope_all_type
slope_all_type = coef(CO2_model_add)[2]

plot(CO2 ~ horse, data = epa2017, col = c(1,4,6), pch = c(1,4,6), cex = 2)
legend("topright", c("Both", "Car", "Truck"), col = c(1,4,6), pch = c(1,4,6))
abline(int_both, slope_all_type, col = 1, pch = 1)
abline(int_car, slope_all_type, col = 4, pch = 4)
abline(int_truck, slope_all_type, col = 6, pch = 6)

```

```{r, echo=FALSE}
summary(CO2_model_add)

predict(CO2_model_add, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)
```
**Comment**
- This 3 regression lines seems to fit the model better than the slr's regression line and we can see the line for each type of car. They are sharing same slops therefore they appear to have 3 regression lines. Based on the plot above, we could also see that $Truck$ > $both$ > $Car$ in term of producing more Co2 which makes sense

- `r coef(CO2_model_add)[2]` is the estimate for the average change in `CO2` for a one foot-pound per second increase in `horse` for a vehicle of type `car`. 

- The 90% prediction interval using this model for the `CO2` of a Subaru Impreza Wagon, which is a vehicle with 148 horsepower and is considered type `Both` has lower bound as `r predict(CO2_model_add, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)[2]` and upper bound as `r predict(CO2_model_add, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)[3]`. 

**(c)** Do the following:

- Make a scatterplot of `CO2` versus `horse`. Use a different color point for each vehicle `type`. 
- Fit an interaction multiple regression model with `CO2` as the response and `horse` and `type` as the predictors.
- Add the fitted regression "lines" to the scatterplot with the same colors as their respective points (one line for each vehicle type).

```{r fig.height=10, fig.width=10, echo=FALSE}
CO2_model_int = lm(CO2 ~ horse * type, data = epa2017) 

#intercept
int_both = coef(CO2_model_int)[1]
int_car = coef(CO2_model_int)[1] + coef(CO2_model_int)[3]
int_truck = coef(CO2_model_int)[1] + coef(CO2_model_int)[4]

#slope_all_type
slope_both = coef(CO2_model_int)[2]
slope_car = coef(CO2_model_int)[2] + coef(CO2_model_int)[5]
slope_truck = coef(CO2_model_int)[2] + coef(CO2_model_int)[6]

plot(CO2 ~ horse, data = epa2017, col = c(1,4,6), pch = c(1,4,6), cex = 2)
legend("topright", c("Both", "Car", "Truck"), col = c(1,4,6), pch = c(1,4,6))
abline(int_both, slope_all_type, col = 1, pch = 1)
abline(int_car, slope_car, col = 4, pch = 4)
abline(int_truck, slope_truck, col = 6, pch = 6)
```

```{r,echo=FALSE}
summary(CO2_model_int)
predict(CO2_model_int, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)

```

**Comment**

- This model gives us 3 regression lines with 3 different slope for types of cars. These regressin line seems to fit the model better than two previous ones and show us better relationship of $horsepower$ and $CO2$ in term of different $type$ of car. We can still see that $CO2$ produce by $truck$ is still the highest, to $both$ then to $car$. However we can see that at the lower $housepower$, the type of $CO2$ produced was about the same in type "car" and type "both". The more $horsepower$ is added the further 2 regression lines to each other. Same for Truck type vs Both type. 

- `r slope_car` Give an estimate for the average change in `CO2` for a one foot-pound per second increase in `horse` for a vehicle of type `car`. 

- The 90% prediction interval using this model for the `CO2` of a Subaru Impreza Wagon, which is a vehicle with 148 horsepower and is considered type `Both` has lower bound as `r predict(CO2_model_int, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)[2]` and upper bound as `r predict(CO2_model_int, data.frame(horse = c(148), type = "Both"), interval = "prediction", level = 0.90)[3]`

**(d)** Based on the previous plots, you probably already have an opinion on the best model. Now use an ANOVA $F$-test to compare the additive and interaction models. Based on this test and a significance level of $\alpha = 0.10$, which model is preferred?

```{r}
anova(CO2_model_add, CO2_model_int)
```

**Comment**
Given that p-value = `r anova(CO2_model_add, CO2_model_int)[2,6]` is lower compare to $\alpha = 0.10$, it's appeared that *Interaction Model* is the best model.

***

## Exercise 2 (Hospital SUPPORT Data, White Blood Cells)

For this exercise, we will use the data stored in [`hospital.csv`](hospital.csv). It contains a random sample of 580 seriously ill hospitalized patients from a famous study called "SUPPORT" (Study to Understand Prognoses Preferences Outcomes and Risks of Treatment). As the name suggests, the purpose of the study was to determine what factors affected or predicted outcomes, such as how long a patient remained in the hospital. The variables in the dataset are:  
 
- `Days` - Days to death or hospital discharge
- `Age` - Age on day of hospital admission
- `Sex` - Female or male
- `Comorbidity` - Patient diagnosed with more than one chronic disease
- `EdYears` - Years of education
- `Education` - Education level; high or low
- `Income` - Income level; high or low
- `Charges` - Hospital charges, in dollars
- `Care` - Level of care required; high or low
- `Race` - Non-white or white
- `Pressure` - Blood pressure, in mmHg
- `Blood` - White blood cell count, in gm/dL
- `Rate` - Heart rate, in bpm

For this exercise, we will use `Age`, `Education`, `Income`, and `Sex` in an attempt to model `Blood`. Essentially, we are attempting to model white blood cell count using only demographic information.


**(a)** Load the data, and check its structure using `str()`. Verify that `Education`, `Income`, and `Sex` are factors; if not, coerce them to be factors. What are the levels of `Education`, `Income`, and `Sex`?

Loading Data

```{r}
hospital <- read.csv("hospital.csv")
```

Checking structure of data

```{r}
str(hospital)
```

Verifying Data
```{r}
is.factor(hospital$Education)

is.factor(hospital$Income)

is.factor(hospital$Sex)
```
Checking levels of factor variables

```{r}
levels(hospital$Education)
```

```{r}
levels(hospital$Income)
```

```{r}
levels(hospital$Sex)
```

**(b)** Fit an additive multiple regression model with `Blood` as the response using `Age`, `Education`, `Income`, and `Sex` as predictors. What does `R` choose as the reference level for `Education`, `Income`, and `Sex`?

```{r}
blood_mol_add = lm(Blood ~ Age + Education + Income + Sex, data = hospital)
summary(blood_mol_add)
```
**Comment**
R is using
- Education: *low* 
- Income: *low* 
_ Sex: *male* 

**(c)** Fit a multiple regression model with `Blood` as the response. Use the main effects of `Age`, `Education`, `Income`, and `Sex`, as well as the interaction of `Sex` with `Age` and the interaction of `Sex` and `Income`. Use a statistical test to compare this model to the additive model using a significance level of $\alpha = 0.10$. Which do you prefer?

Creating interaction select model
```{r}
blood_mol_int2 = lm(Blood ~ Age + Education + Income + Sex + Sex:Age + Sex:Income, data = hospital)
anova(blood_mol_add,blood_mol_int2)

```

Since p-value of AnovaFstatistic is `r anova(blood_mol_add,blood_mol_int2)[2,6]` > than $\alpha = 0.10$, therefore I prefered additive more.

**(d)** Fit a model similar to that in **(c)**, but additionally add the interaction between `Income` and `Age` as well as a three-way interaction between `Age`, `Income`, and `Sex`. Use a statistical test to compare this model to the preferred model from

```{r}
blood_mol_int3 = lm(Blood ~ Age + Education + Income + Sex + Sex:Age + Sex:Income + Income:Age + Age:Income:Sex, data = hospital)
anova(blood_mol_add,blood_mol_int3)
```

Since p-value of AnovaFstatistic is `r anova(blood_mol_add,blood_mol_int3)[2,6]` < than $\alpha = 0.10$, I prefered the new model with three-way interaction more.

**(e)** Using the model in **(d)**, give an estimate of the change in average `Blood` for a one-unit increase in `Age` for a highly educated, low income, male patient.

```{r,echo=FALSE}
avg_change = coef(blood_mol_int3)["Age"] + coef(blood_mol_int3)["Age:Sexmale"] + coef(blood_mol_int3)["Age:Incomelow"] + coef(blood_mol_int3)["Age:Incomelow:Sexmale"]
```

The estimate of the change in average `Blood` for a one-unit increase in `Age` for a highly educated, low income, male patient is `r avg_change`
***

## Exercise 3 (Hospital SUPPORT Data, Stay Duration)

For this exercise, we will again use the data stored in [`hospital.csv`](hospital.csv). It contains a random sample of 580 seriously ill hospitalized patients from a famous study called "SUPPORT" (Study to Understand Prognoses Preferences Outcomes and Risks of Treatment). As the name suggests, the purpose of the study was to determine what factors affected or predicted outcomes, such as how long a patient remained in the hospital. The variables in the dataset are:  
 
- `Days` - Days to death or hospital discharge
- `Age` - Age on day of hospital admission
- `Sex` - Female or male
- `Comorbidity` - Patient diagnosed with more than one chronic disease
- `EdYears` - Years of education
- `Education` - Education level; high or low
- `Income` - Income level; high or low
- `Charges` - Hospital charges, in dollars
- `Care` - Level of care required; high or low
- `Race` - Non-white or white
- `Pressure` - Blood pressure, in mmHg
- `Blood` - White blood cell count, in gm/dL
- `Rate` - Heart rate, in bpm

For this exercise, we will use `Blood`, `Pressure`, and `Rate` in an attempt to model `Days`. Essentially, we are attempting to model the time spent in the hospital using only health metrics measured at the hospital.

Consider the model

\[
Y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3 + \beta_4 x_1 x_2 + \beta_5 x_1 x_3 + \beta_6 x_2 x_3 + \beta_7 x_1 x_2 x_3 + \epsilon,
\]

where

- $Y$ is `Days`
- $x_1$ is `Blood`
- $x_2$ is `Pressure`
- $x_3$ is `Rate`.

**(a)** Fit the model above. Also fit a smaller model using the provided `R` code.

```{r}
days_add = lm(Days ~ Pressure + Blood + Rate, data = hospital)
days_int = lm(Days ~ (Pressure + Blood + Rate)^3, data = hospital)

anova(days_add,days_int)
```

Use a statistical test to compare the two models. Report the following:

- The null and alternative hypotheses in terms of the model given in the exercise description
- The value of the test statistic
- The p-value of the test
- A statistical decision using a significance level of $\alpha = 0.10$
- Which model you prefer

**Answer**

- $H_0 = \beta_4 = \beta_5 = \beta_6 = \beta_7 = 0$ 
- Test statistic is `r anova(days_add,days_int)$F[2]`
- p-value : `r anova(days_add,days_int)[2,"Pr(>F)"]`
- Based on p-value provided, I will pick interaction model as my preferred model

**(b)** Give an expression based on the model in the exercise description for the true change in length of hospital stay in days for a 1 bpm increase in `Rate` for a patient with a `Pressure` of 139 mmHg and a `Blood` of 10 gm/dL. Your answer should be a linear function of the $\beta$s.

Full formula

\[
Y = \beta_0 + 10\beta_1 + 139\beta_2 + 1390\beta_4 + (\beta_3 + 10\beta_5 + 139\beta_6* + 1390\beta_7)x_3 + e,
\]

True change for a 1 bpm increase in `Rate` for a patient with a `Pressure` of 139 mmHg and a `Blood` of 10 gm/dL. Your answer should be a linear function of the $\beta$s.

\[
Y = \beta_3 + 10\beta_5 + 139\beta_6* + 1390\beta_7
\]

**(c)** Give an expression based on the additive model in part **(a)** for the true change in length of hospital stay in days for a 1 bpm increase in `Rate` for a patient with a `Pressure` of 139 mmHg and a `Blood` of 10 gm/dL. Your answer should be a linear function of the $\beta$s.

$Y = \beta_0 + 10\beta_1 + 139\beta_2 + \beta_3 + e$

True change:

$Y = \beta_3$

***

## Exercise 4 ($t$-test Is a Linear Model)

In this exercise, we will try to convince ourselves that a two-sample $t$-test assuming equal variance is the same as a $t$-test for the coefficient in front of a single two-level factor variable (dummy variable) in a linear model.

First, we set up the data frame that we will use throughout.

```{r}
n = 30

sim_data = data.frame(
  groups = c(rep("A", n / 2), rep("B", n / 2)),
  values = rep(0, n))
str(sim_data)
```

We will use a total sample size of `30`, `15` for each group. The `groups` variable splits the data into two groups, `A` and `B`, which will be the grouping variable for the $t$-test and a factor variable in a regression. The `values` variable will store simulated data.

We will repeat the following process a number of times.

```{r}
set.seed(20)
sim_data$values = rnorm(n, mean = 42, sd = 3.5) # simulate response data
summary(lm(values ~ groups, data = sim_data))
t.test(values ~ groups, data = sim_data, var.equal = TRUE)
```

We use `lm()` to test

\[
H_0: \beta_1 = 0
\]

for the model

\[
Y = \beta_0 + \beta_1 x_1 + \epsilon
\]

where $Y$ is the values of interest, and $x_1$ is a dummy variable that splits the data in two. We will let `R` take care of the dummy variable.

We use `t.test()` to test

\[
H_0: \mu_A = \mu_B
\]

where $\mu_A$ is the mean for the `A` group, and $\mu_B$ is the mean for the `B` group.

The following code sets up some variables for storage.

```{r}
num_sims = 300
lm_t = rep(0, num_sims)
lm_p = rep(0, num_sims)
tt_t = rep(0, num_sims)
tt_p = rep(0, num_sims)
```

- `lm_t` will store the test statistic for the test $H_0: \beta_1 = 0$.
- `lm_p` will store the p-value for the test $H_0: \beta_1 = 0$.
- `tt_t` will store the test statistic for the test $H_0: \mu_A = \mu_B$.
- `tt_p` will store the p-value for the test $H_0: \mu_A = \mu_B$.

The variable `num_sims` controls how many times we will repeat this process, which we have chosen to be `300`.

**(a)** Set a seed equal to your birthday. Then write code that repeats the above process `300` times. Each time, store the appropriate values in `lm_t`, `lm_p`, `tt_t`, and `tt_p`. Specifically, each time you should use `sim_data$values = rnorm(n, mean = 42, sd = 3.5)` to update the data. The grouping will always stay the same.

```{r}
set.seed(19920531)

for (i in 1:num_sims){
  sim_data$values = rnorm(n, mean = 42, sd = 3.5) # simulate response data
  summary_model = summary(lm(values ~ groups, data = sim_data))
  test = t.test(values ~ groups, data = sim_data, var.equal = TRUE)
  
  #storing value
  lm_t[i] = coef(summary_model)[2,"t value"]
  lm_p[i] = coef(summary_model)[2,"Pr(>|t|)"]
  tt_t[i] = test$statistic
  tt_p[i] = test$p.value
}

```


**(b)** Report the value obtained by running `mean(lm_t == tt_t)`, which tells us what proportion of the test statistics is equal. The result may be extremely surprising!

```{r}
mean(lm_t == tt_t)
```

The proportion being equal is `r mean(lm_t == tt_t)`

**(c)** Report the value obtained by running `mean(lm_p == tt_p)`, which tells us what proportion of the p-values is equal. The result may be extremely surprising!

```{r}
mean(lm_p == tt_p)
```


**(d)** If you have done everything correctly so far, your answers to the last two parts won't indicate the equivalence we want to show! What the heck is going on here? The first issue is one of using a computer to do calculations. When a computer checks for equality, it demands **equality**; nothing can be different. However, when a computer performs calculations, it can only do so with a certain level of precision. So, if we calculate two quantities we know to be analytically equal, they can differ numerically. Instead of `mean(lm_p == tt_p)` run `all.equal(lm_p, tt_p)`. This will perform a similar calculation, but with a very small error tolerance for each equality. What is the result of running this code? What does it mean?

```{r}
all.equal(lm_p, tt_p)
```
Base on the result of this code as p-value for the test $H_0: \beta_1 = 0$ and p-value for the test  $H_0: \mu_A = \mu_B$ is equal. This means that even though p value might be different. Those differences are small and can be denoted to error. They are analyticaly equall.

**(e)** Your answer in **(d)** should now make much more sense. Then what is going on with the test statistics? Look at the values stored in `lm_t` and `tt_t`. What do you notice? Is there a relationship between the two? Can you explain why this is happening?

Looking at the the values in `lm_t` and `tt_t`, I noticed that the value for both group in each simulation is the same but signed opposite of eachother. This can be explained by equation of t-test in linear model and two equation t-test. Therefore it will be depending $beta_1$ to estimate the sign of specific equation

$$t = \beta_1 - 0$$
$$t = \mu_A - \mu_B - 0$$


