---
title: "Week 1 - Homework"
author: "STAT 420, Summer 2020, D. Unger"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---


***

## Exercise 1 (Subsetting and Statistics)

For this exercise, we will use the `msleep` dataset from the `ggplot2` package.

**(a)** Install and load the `ggplot2` package. **Do not** include the installation command in your `.Rmd` file. (If you do it will install the package every time you knit your file.) **Do** include the command to load the package into your environment.

```{r not_run, eval = FALSE, echo=FALSE}
install.packages(ggplot2)
```

```{r}
library(ggplot2)
library(tibble)
msleep = ggplot2::msleep
msleep = as_tibble(msleep)
```


**(b)** Note that this dataset is technically a `tibble`, not a data frame. How many observations are in this dataset? How many variables? What are the observations in this dataset?
**Answer**
There are:
- `r nrow(msleep)` observations 
- `r ncol(msleep)` variables

The observations include:
```{r, echo = FALSE }
as_tibble(msleep)
```

**(c)** Mean hours of REM sleep of individuals in this dataset is **`r mean(msleep$sleep_rem, na.rm = TRUE)`**

**(d)** The standard deviation of brain weight of individuals in this dataset is **`r sd(msleep$brainwt, na.rm = TRUE)`**

**(e)** **`r msleep[which.max(msleep$sleep_rem),]$name `** in this dataset gets the most REM sleep

**(f)** The average bodyweight of carnivores in this dataset is **`r mean(msleep[msleep$vore == "carni",]$bodywt, na.rm = TRUE)`**

***

## Exercise 2 (Plotting)

For this exercise, we will use the `birthwt` dataset from the `MASS` package.

**(a)** Note that this dataset is a data frame and all of the variables are numeric. How many observations are in this dataset? How many variables? What are the observations in this dataset?
```{r , eval = FALSE, echo=FALSE}
if (!require("MASS")) {
  install.packages("MASS") 
  require(MASS)
}

```

```{r}
birthwt = MASS::birthwt
birthwt = as_tibble(birthwt)
```

There are:
- **`r nrow(birthwt)`** observations 
- **`r ncol(birthwt)`** variables

The observations include:
```{r, echo = FALSE }
as_tibble(birthwt)
```


**(b)** Below is a scatter plot of birth weight (y-axis) vs mother's weight before pregnancy (x-axis) 
```{r, echo = FALSE}

plot(x = birthwt$lwt, y = birthwt$bwt,
     xlab ="Mother's weight before pregnancy",
     ylab = "Birth weight",
     main = "Relationship between infant's weight vs mother's weight",
     col = "darkorange",
     ) 
```

**Comment:** Based on the scatter plot above, we can see there is a very week linear rel = 

**(c)** Create a scatter plot of birth weight (y-axis) vs mother's age (x-axis). Use a non-default color for the points. (Also, be sure to give the plot a title and label the axes appropriately.) Based on the scatter plot, does there seem to be a relationship between the two variables? Briefly explain. 

```{r,echo=FALSE}
plot(x = birthwt$age, y = birthwt$bwt,
     xlab = "Mother's age",
     ylab = "Birth weight",
     main = "Relationship between infant's weight vs mother's age",
     col = "blue",
     ) 
```

**Comment** The data is well spread and is almost flat suggesting that there is not much of a relationship between the mother's age and infant's birthweight

**(d)** Create side-by-side boxplots for birth weight grouped by smoking status. Use non-default colors for the plot. (Also, be sure to give the plot a title and label the axes appropriately.) Based on the boxplot, does there seem to be a difference in birth weight for mothers who smoked? Briefly explain.


```{r,echo=FALSE}
boxplot (bwt ~ smoke, data = birthwt,
         xlab = "Smoke Status. 0 = No, 1= yes ",
         ylab = " Birth Weight",
         main = "Birthwt vs Smoking Status",
         col = "lightblue",
         border = "navy"
         )
```

**Comment** Comparing to box size and means, we can see that smoking might be a factor of low-weight babies. Women who smoke tend to have lighter/smaller comparing to non-smoker women. In the "Smowking" box there is also data that show an outliner with baby weighted less than 1000gram.

***

## Exercise 3 (Importing Data, More Plotting)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` - in kcal
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

```{r , echo=FALSE}
nutrition.2018 <- read.csv("nutrition-2018.csv")
```

**(a)** Create a histogram of `Calories`. Do not modify `R`'s default bin selection. Make the plot presentable. Describe the shape of the histogram. Do you notice anything unusual?
```{r, echo=FALSE}
hist(nutrition.2018$Calories,
     plot = TRUE,
     xlab = "Calories (kcal)",
     main = "Histogram of Calories",
     col = "lightgreen"
     )
```

**Comment** The histogram below is right-skewed shave meaning that most of the food are sized in the lower scale of calories and are kept under 900. There are some odd spikes in the histogram, one in 400 kcals and one in 800 kcal. 

**(b)** Create a scatter plot of calories (y-axis) vs protein (x-axis). Make the plot presentable. Do you notice any trends? Do you think that knowing only the protein content of a food, you could make a good prediction of the calories in the food?

```{r, echo=FALSE}
plot(x = nutrition.2018$Protein, y = nutrition.2018$Calories,
     xlab = "Protein",
     ylab = "Calories",
     main = "Relationship between Protein vs Calories",
     col = "blue",
     ) 
```

**Comment** Protein and Calories seems to have a positive associate from 0 - 15 in protein and 0- 500 in Calories. From 20 grams in Protein we can see the downward slope showing the negative association. Since it seem to have strong relationship between 2 variables I believe we can somewhat predict the calories of the food given protein

**(c)** Create a scatter plot of `Calories` (y-axis) vs `4 * Protein + 4 * Carbs + 9 * Fat` (x-axis). Make the plot presentable. You will either need to add a new variable to the data frame, or use the `I()` function in your formula in the call to `plot()`. If you are at all familiar with nutrition, you may realize that this formula calculates the calorie count based on the protein, carbohydrate, and fat values. You'd expect then that the result here is a straight line. Is it? If not, can you think of any reasons why it is not?

```{r, echo=FALSE}

plot(nutrition.2018$Calories ~ I(4 * nutrition.2018$Protein + 4 * nutrition.2018$Carbs + 9 * nutrition.2018$Fat),
     xlab = "Calcount based on protein, carb + fat",
     ylab = "Calories (in Kcal)",
     main = " Calcount (protein + carbohydrate + fat vs Calories",
     col = "yellow",
     ) 

```

**Comment**
It is not perfectly straightly since there are some outliners could be caused by some rounding? or methodology?
***

## Exercise 4 (Writing and Using Functions)

For each of the following parts, use the following vectors:

```{r}
a = 1:10
b = 10:1
c = rep(1, times = 10)
d = 2 ^ (1:10)
```

**(a)** Write a function called `sum_of_squares`.

- Arguments:
    - A vector of numeric data `x`
- Output:
    - The sum of the squares of the elements of the vector $\sum_{i = 1}^n x_i^2$
    
Provide your function, as well as the result of running the following code:

```{r}
sum_of_squares <- function(x){
  sum(x^2)
}

sum_of_squares(x = a)
sum_of_squares(x = c(c, d))
```

**(b)** Using only your function `sum_of_squares()`, `mean()`, `sqrt()`, and basic math operations such as `+` and `-`, calculate

\[
\sqrt{\frac{1}{n}\sum_{i = 1}^n (x_i - 0)^{2}}
\]

where the $x$ vector is `d`.
```{r}
sqrt(sum_of_squares(x = d)/length(d))
```


**(c)** Using only your function `sum_of_squares()`, `mean()`, `sqrt()`, and basic math operations such as `+` and `-`, calculate

\[
\sqrt{\frac{1}{n}\sum_{i = 1}^n (x_i - y_i)^{2}}
\]

where the $x$ vector is `a` and the $y$ vector is `b`.
```{r}
sqrt(sum_of_squares(x = a - b)/length(a))
```


***

## Exercise 5 (More Writing and Using Functions)

For each of the following parts, use the following vectors:

```{r}
set.seed(42)
x = 1:100
y = rnorm(1000)
z = runif(150, min = 0, max = 1)
```

**(a)** Write a function called `list_extreme_values`.

- Arguments:
    - A vector of numeric data `x`
    - A positive constant, `k`, with a default value of `2`
- Output:
    - A list with two elements:
        - `small`, a vector of elements of `x` that are $k$ sample standard deviations less than the sample mean. That is, the observations that are smaller than $\bar{x} - k \cdot s$.
        - `large`, a vector of elements of `x` that are $k$ sample standard deviations greater than the sample mean. That is, the observations that are larger than $\bar{x} + k \cdot s$.

Provide your function, as well as the result of running the following code:

```{r}
list_extreme_values = function(x,k){
  smallVal = mean(x)- k * sd(x)
  largeVal = mean(x)+ k * sd(x)
  list ((x[x<smallVal]),(x[x>largeVal]))
}

list_extreme_values(x = x, k = 1)
list_extreme_values(x = y, k = 3)
list_extreme_values(x = y, k = 2)
list_extreme_values(x = z, k = 1.5)
```

**(b)** Using only your function `list_extreme_values()`, `mean()`, and basic list operations, calculate the mean of observations that are greater than 1.5 standard deviation above the mean in the vector `y`.
```{r}
mean(list_extreme_values(x = y, k = 1.5)[[2]])

```

