---
title: "FinalProject (1)"
author: "Anh Nguyen"
date: "8/5/2020"
output: html_document
---
#Data Cleaning
```{r}
library(faraway)
```


```{r}
data = read.csv("kc_house_data.csv")
all.equal(data, na.omit(data))
```

```{r}
month = as.factor(substr(data$date, 5, 6))
age = as.integer(substr(data$date, 1, 4)) - ifelse(data$yr_renovated == 0, as.integer(data$yr_built), as.integer(data$yr_renovated))
data = subset(data, select = -c(id, lat, long, date, yr_built, yr_renovated))
data$waterfront = as.factor(data$waterfront)
data$view = as.factor(data$view)
data$condition = as.factor(data$condition)
data$grade = as.factor(data$grade)
data$floors = as.factor(data$floors)
data$bedrooms = as.factor(data$bedrooms)
data$bathrooms = as.factor(data$bathrooms)
data$zipcode = as.factor(data$zipcode)
data = cbind(data, month, age)
```

Function

```{r}
get_metrics = function(model) {
  metrics = c(sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2)),
              summary(model)$sigma,
              summary(model)$adj.r.squared)
  names(metrics) = c("LOOCV RMSE", "RSE", "Adj. R Sqr")
  metrics
}

#drawing plot
draw_graph = function (model, pcol = "gray", lcol = "dodgerblue") {
    par(mfrow = c(1,2))
    #fitted vs residual plots
    plot(fitted(model), resid(model), col = "grey", pch = 20,
      xlab = "Fitted", ylab = "Residuals", main = "Fitted vs Residual Graph")
      abline(h = 0, col = "darkorange", lwd = 2)
      
    #QQ Plot 
    qqnorm(resid(model), main = "Normal Q-Q Plot" , col = pcol)
    qqline(resid(model), col = lcol, lwd = 2)
}
```

Among 7 numeric variables we are now choose to use $sqft_living$ & $sqft_living15$ since they are both highly corelated to $price$ and not as correlated to each other. Also one of the curiosities that we have while doing this project was whether differnt zipcode drives different price range for Seattle Housing, we have added also add interaction between 2 aboves variable with $Zipcode$
```{r} 
int_mod = lm(price ~ sqft_living +  sqft_living15 + zipcode + sqft_living:zipcode + sqft_living15:zipcode, data = data)
get_metrics(int_mod)
```

```{r}
draw_graph(int_mod)
```

Since the RMSE is high . Using log transformation for numeric variables + predictor
```{r}
#int_mod_1 = lm (log(price) ~ . - sqft_living - sqft_above + log(sqft_living) + I(sqft_living^2)  + sqft_living:zipcode , data = data)
int_mod_1 = lm (log(price) ~  log(sqft_living15) + log(sqft_living) + zipcode  + sqft_living:zipcode + sqft_living15:zipcode, data = data)
get_metrics(int_mod_1)
```

```{r}

draw_graph(int_mod_1)
```


```{r}
int_mod_2 = lm (log(price) ~  sqft_living15 + I(sqft_living15^2)  + sqft_living + I(sqft_living^2)+ zipcode  + sqft_living:zipcode + sqft_living15:zipcode, data = data)
get_metrics(int_mod_2)
```

```{r}
draw_graph(int_mod_2)
```
 
```{r}

outliner = as.vector(as.integer(names(rstandard(int_mod_1)[abs(rstandard(int_mod_1)) > 2])))
high_influence = as.vector(which(cooks.distance(int_mod_1)> 4/ length(cooks.distance(int_mod_1))))
tobeRemove = c(outliner,high_influence)
new_data = data[-tobeRemove,]

```

```{r}
new_int_mod = lm (log(price) ~  log(sqft_living15) + log(sqft_living) + zipcode  + sqft_living:zipcode + sqft_living15:zipcode, data = new_data)
get_metrics(new_int_mod)
```

```{r}
draw_graph(new_int_mod)
```

