---
title: "Week 2 - Homework"
author: "STAT 420, Summer 2020, D. Unger"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes  
  pdf_document: default
urlcolor: cyan
---

***

## Exercise 1 (Using `lm`)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Suppose we would like to understand the size of a cat's heart based on the body weight of a cat. Fit a simple linear model in `R` that accomplishes this task. Store the results in a variable called `cat_model`. Output the result of calling `summary()` on `cat_model`.

```{r, echo=FALSE}
cats = MASS :: cats

cat_model = lm(Hwt ~ Bwt, data = cats)
```

**Output**

```{r}
summary(cat_model)
```


**(b)** Output only the estimated regression coefficients. Interpret $\hat{\beta_0}$ and $\beta_1$ in the *context of the problem*. Be aware that only one of those is an estimate.

```{r}
coef(cat_model)
```

**Comment**
- $\beta_1$ tell us that for the increse of cat's bodyweight in kg, there is also the mean increasing of cat's heart weight in g
- The intercept parameter $\hat{\beta_0} `r coef(cat_model)[1]`tells us the mean of cat's hearth weight when cat's bodyweight is 0*

**(c)** Use your model to predict the heart weight of a cat that weights **3.1** kg. Do you feel confident in this prediction? Briefly explain.
```{r}
predict(cat_model, newdata = data.frame(Bwt = 3.1))
```
```{r}
unique(cats$Bwt)

3.1 %in% unique(cats$Bwt)
```

**Answer**: I feel confident about this prediction since this weight is an observed value of x (inside of the data range).

**(d)** Use your model to predict the heart weight of a cat that weights **1.5** kg. Do you feel confident in this prediction? Briefly explain.
```{r}
predict(cat_model, newdata = data.frame(Bwt = 1.5))
```

```{r}
unique(cats$Bwt)

1.5 %in% unique(cats$Bwt)
```

**Answer**: Since this data is not an observed value of x (also outside data range), I'm less confident in prediction of this

**(e)** Create a scatterplot of the data and add the fitted regression line. Make sure your plot is well labeled and is somewhat visually appealing.

```{r}
plot(Hwt ~ Bwt, data = cats,
     xlab = "Bodyweight (in kg)",
     ylab = "Heartweight (in g)",
     main = "Cat's Heart Weight vs Body Weight",
     pch = 20,
     cex = 2,
     col = "grey"
     )

abline(cat_model, lwd =3, col = "darkorange")
```

**(f)** Report the value of $R^2$ for the model. Do so directly. Do not simply copy and paste the value from the full output in the console after running `summary()` in part **(a)**.

```{r}
x = cats$Bwt
y = cats$Hwt

Sxy = sum((x - mean(x)) * (y - mean(y)))
Sxx = sum((x - mean(x)) ^ 2)
Syy = sum((y - mean(y)) ^ 2)

beta_1_hat = Sxy / Sxx
beta_0_hat = mean(y) - beta_1_hat * mean(x)

y_hat = beta_0_hat + beta_1_hat * x

SST   = sum((y - mean(y)) ^ 2)
SSReg = sum((y_hat - mean(y)) ^ 2)
SSE   = sum((y - y_hat) ^ 2)
R2 = SSReg / SST

R2
```

**Answer**: Value of $R^2$ is `r summary(cat_model)$r.squared`

****

## Exercise 2 (Writing Functions)

This exercise is a continuation of Exercise 1.

**(a)** Write a function called `get_sd_est` that calculates an estimate of $\sigma$ in one of two ways depending on input to the function. The function should take three arguments as input:

- `fitted_vals` - A vector of fitted values from a model
- `actual_vals` - A vector of the true values of the response
- `mle` - A logical (`TRUE` / `FALSE`) variable which defaults to `FALSE`

The function should return a single value:

- $s_e$ if `mle` is set to `FALSE`.
- $\hat{\sigma}$ if `mle` is set to `TRUE`.

```{r}
get_sd_est = function (fitted_vals, actual_vals, mle = FALSE){
  e = actual_vals - fitted_vals
  n = length(e)
  return(
    if(mle == TRUE) sqrt(sum(e^2)/n)
    else sqrt(sum(e^2)/(n-2))
  )
}
```

**(b)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `FALSE`. Explain the resulting estimate in the context of the model.

```{r}
get_sd_est(fitted.values(cat_model),cats$Hwt, mle = FALSE)
```

**Explain**: Given Least Square Model'sestimated value of from catmodel & actual value from cat's heartweight where we have adjust degree of freedom of 2. we have standard deviation for cat's heartweight as `r get_sd_est(fitted.values(cat_model),cats$Hwt, mle = FALSE)` which is closer to the summary.

**(c)** Run the function `get_sd_est` on the residuals from the model in Exercise 1, with `mle` set to `TRUE`. Explain the resulting estimate in the context of the model. Note that we are trying to estimate the same parameter as in part **(b)**.

```{r}
get_sd_est(fitted.values(cat_model),cats$Hwt, mle = TRUE)
```

**Explain** Given Maximum Likelihood estimated value of from catmodel & actual value from cat's heartweight where we DON'T adjust degree of freedom. we have standard deviation for cat's heartweight as `r get_sd_est(fitted.values(cat_model),cats$Hwt, mle = TRUE)`.


**(d)** To check your work, output `summary(cat_model)$sigma`. It should match at least one of **(b)** or **(c)**.
```{r}
summary(cat_model)$sigma
```

***

## Exercise 3 (Simulating SLR)

Consider the model

\[
Y_i = 5 + -3 x_i + \epsilon_i
\]

with 

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 10.24)
\]

where $\beta_0 = 5$ and $\beta_1 = -3$.

This exercise relies heavily on generating random observations. To make this reproducible we will set a seed for the randomization. Alter the following code to make `birthday` store your birthday in the format: `yyyymmdd`. For example, [William Gosset](https://en.wikipedia.org/wiki/William_Sealy_Gosset), better known as *Student*, was born on June 13, 1876, so he would use:

```{r}
birthday = 19920531
set.seed(birthday)
```

**(a)** Use `R` to simulate `n = 25` observations from the above model. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
x = runif(n = 25, 0, 10)
```

You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Store the data frame this function returns in a variable of your choice. Note that this function calls $y$ `response` and $x$ `predictor`.

```{r}
sim_slr = function(x, beta_0 = 5, beta_1 = -3, sigma = 1) {
  n = length(x)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  data.frame(predictor = x, response = y)
}
```

```{r}
set.seed(1)
sim_data = sim_slr(x, beta_0 = 5, beta_1 = -3, sigma = sqrt(10.24))

```

**(b)** Fit a model to your simulated data. Report the estimated coefficients. Are they close to what you would expect? Briefly explain.
```{r}
sim_fit = lm(response ~ predictor, data = sim_data)
coef(sim_fit)
```

They are somewhat closed to $\beta_0$ and $\beta_1$ since they are the estimated $\beta_0 = 5$ and $\beta_1$ using  generated observations

**(c)** Plot the data you simulated in part **(a)**. Add the regression line from part **(b)** as well as the line for the true model. Hint: Keep all plotting commands in the same chunk.

```{r}
plot(response ~ predictor, data = sim_data,
     xlab = "Simulated Predictor Variable",
     ylab = "Simulated Response Variable",
     main = "Simulated Regression Data",
     pch  = 20,
     cex  = 2,
     col  = "grey")

abline(sim_fit, lwd = 3, lty = 1, col = "darkorange")
abline(5, -3, lty = 2, col = "dodgerblue")
legend("topright", c("Estimate", "Truth"), lty = c(1, 2), lwd = 2,
       col = c("darkorange", "dodgerblue"))

```


**(d)** Use `R` to repeat the process of simulating `n = 25` observations from the above model $1500$ times. Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. Some hints:

```{r}
beta_hat_1 = rep(0,1500)

for (i in 1:1500) {
  sim_data = sim_slr(x = x, beta_0 = 5, beta_1 = -3, sigma = 3.2)
  model = lm(response ~ predictor, data = sim_data)
  beta_hat_1[i] = coef(model)[2]
}

```

- Consider a `for` loop.
- Create `beta_hat_1` before writing the `for` loop. Make it a vector of length $1500$ where each element is `0`.
- Inside the body of the `for` loop, simulate new $y$ data each time. Use a variable to temporarily store this data together with the known $x$ data as a data frame.
- After simulating the data, use `lm()` to fit a regression. Use a variable to temporarily store this output.
- Use the `coef()` function and `[]` to extract the correct estimated coefficient.
- Use `beta_hat_1[i]` to store in elements of `beta_hat_1`.
- See the notes on [Distribution of a Sample Mean](http://daviddalpiaz.github.io/appliedstats/introduction-to-r.html#distribution-of-a-sample-mean) for some inspiration.

You can do this differently if you like. Use of these hints is not required.

**(e)** Report the mean and standard deviation of `beta_hat_1`. Do either of these look familiar?
```{r}
mean(beta_hat_1)
sd(beta_hat_1)
```
**Comment**: the mean of beta_hat_1 looks familier, very close to $\beta_1$ = -3

**(f)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.
```{r}
hist(beta_hat_1,
     plot = TRUE,
     col = "lightgreen")
```
**Comment** The shape is Normal Distribution

***

## Exercise 4 (Be a Skeptic)

Consider the model

\[
Y_i = 3 + 0 \cdot x_i + \epsilon_i
\]

with

\[
\epsilon_i \sim N(\mu = 0, \sigma^2 = 4)
\]

where $\beta_0 = 3$ and $\beta_1 = 0$.

Before answering the following parts, set a seed value equal to **your** birthday, as was done in the previous exercise.

```{r}
birthday = 19920531
set.seed(birthday)
```

**(a)** Use `R` to repeat the process of simulating `n = 75` observations from the above model $2500$ times. For the remainder of this exercise, use the following "known" values of $x$.

```{r}
x = runif(n = 75, 0, 10)
```

Each time fit a SLR model to the data and store the value of $\hat{\beta_1}$ in a variable called `beta_hat_1`. You may use [the `sim_slr ` function provided in the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#simulating-slr). Hint: Yes $\beta_1 = 0$.

```{r}
sim_slr = function(x, beta_0 = 5, beta_1 = -3, sigma = 1) {
  n = length(x)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  data.frame(predictor = x, response = y)
}
```

```{r}
beta_hat_1 = rep(0,2500)

for (i in 1:2500) {
  sim_data = sim_slr(x, beta_0 = 3, beta_1 = 0, sigma = 2)
  model = lm(response ~ predictor, data = sim_data)
  beta_hat_1[i] = coef(model)[2]
}
```

**(b)** Plot a histogram of `beta_hat_1`. Comment on the shape of this histogram.

```{r}
hist(beta_hat_1,
     plot = TRUE,
     col = "lightpink")
```
**Comment** This histogram seems like a normal distribution with Mean is closer to 0

**(c)** Import the data in [`skeptic.csv`](skeptic.csv) and fit a SLR model. The variable names in `skeptic.csv` follow the same convention as those returned by `sim_slr()`. Extract the fitted coefficient for $\beta_1$.
```{r}
skeptic = read.csv("skeptic.csv")
skeptic_model = lm(skeptic)
skeptic_betahat_1 = coef(skeptic_model)[2]
```

**(d)** Re-plot the histogram from **(b)**. Now add a vertical red line at the value of $\hat{\beta_1}$ in part **(c)**. To do so, you'll need to use `abline(v = c, col = "red")` where `c` is your value.

```{r}
hist(beta_hat_1,
     plot = TRUE,
     col = "lightpink")
abline(v = skeptic_betahat_1, col = "red")
```

**(e)** Your value of $\hat{\beta_1}$ in **(c)** should be negative. What proportion of the `beta_hat_1` values is smaller than your $\hat{\beta_1}$? Return this proportion, as well as this proportion multiplied by `2`.

```{r}
pro = pnorm (skeptic_betahat_1, mean = mean(beta_hat_1), sd = sd(beta_hat_1))
pro 
pro * 2
```


**(f)** Based on your histogram and part **(e)**, do you think the [`skeptic.csv`](skeptic.csv) data could have been generated by the model given above? Briefly explain.

**Comment** We can generate skeptic by the model given even with a very small properbility.
***

## Exercise 5 (Comparing Models)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

```{r, eval = FALSE}
install.packages("mlbench")
```

For simplicity, we will perform some data cleaning before proceeding.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

We have:

- Loaded the data from the package
- Subset the data to relevant variables
    - This is not really necessary (or perhaps a good idea) but it makes the next step easier
- Given variables useful names
- Removed any observation with missing values
    - This should be given much more thought in practice

For this exercise we will define the "Root Mean Square Error" of a model as

\[
\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i = 1}^{n}(y_i - \hat{y}_i)^2}.
\]

**(a)** Fit three SLR models, each with "ozone" as the response. For the predictor, use "wind speed," "humidity percentage," and "temperature" respectively. For each, calculate $\text{RMSE}$ and $R^2$. Arrange the results in a markdown table, with a row for each model. Suggestion: Create a data frame that stores the results, then investigate the `kable()` function from the `knitr` package.

```{r}
# Creating help formula
get_rmse = function(model){
  sqrt(mean((resid(model))^2))
}

get_r2 = function(model){
  summary(model)$r.squared
}
```

```{r}
# Creating 3 different Models
wind_speed_model = lm(ozone ~ wind, data = Ozone)
hum_model = lm(ozone ~ humidity, data = Ozone)
temp_model = lm(ozone ~ temp, data = Ozone)
list = c(wind_speed_model,hum_model,temp_model)
```

```{r}
#cal RMSE - List pf RMSE
RMSE = c(get_rmse(wind_speed_model), get_rmse(hum_model),get_rmse(temp_model)) 

R2 = c(get_r2(wind_speed_model), get_r2(hum_model),get_r2(temp_model)) 
```

```{r}
#create dataframe
result = data.frame(
  Model = c("wind","humidity","temp"),
  RMSE = RMSE,
  R2 = R2
)
```

```{r}
#create markdown table of result
knitr::kable(result)
```


**(b)** Based on the results, which of the three predictors used is most helpful for predicting ozone readings? Briefly explain.

**Comment** Based on the result, the 3rd predictor (Temp) seems to be the most helpful one since they have the lowest RMSE.
***

## Exercise 00 (SLR without Intercept)

**This exercise will _not_ be graded and is simply provided for your information. No credit will be given for the completion of this exercise. Give it a try now, and be sure to read the solutions later.**

Sometimes it can be reasonable to assume that $\beta_0$ should be 0. That is, the line should pass through the point $(0, 0)$. For example, if a car is traveling 0 miles per hour, its stopping distance should be 0! (Unlike what we saw in the book.)

We can simply define a model without an intercept,

\[
Y_i = \beta x_i + \epsilon_i.
\]

**(a)** [In the **Least Squares Approach** section of the text](http://daviddalpiaz.github.io/appliedstats/simple-linear-regression.html#least-squares-approach) you saw the calculus behind the derivation of the regression estimates, and then we performed the calculation for the `cars` dataset using `R`. Here you need to do, but not show, the derivation for the slope only model. You should then use that derivation of $\hat{\beta}$ to write a function that performs the calculation for the estimate you derived. 

In summary, use the method of least squares to derive an estimate for $\beta$ using data points $(x_i, y_i)$ for $i = 1, 2, \ldots n$. Simply put, find the value of $\beta$ to minimize the function

\[
f(\beta)=\sum_{i=1}^{n}(y_{i}-\beta x_{i})^{2}.
\]

Then, write a function `get_beta_no_int` that takes input:

- `x` - A predictor variable
- `y` - A response variable

The function should then output the $\hat{\beta}$ you derived for a given set of data.

**(b)** Write your derivation in your `.Rmd` file using TeX. Or write your derivation by hand, scan or photograph your work, and insert it into the `.Rmd` as an image. See the [RMarkdown documentation](http://rmarkdown.rstudio.com/) for working with images.

**(c)** Test your function on the `cats` data using body weight as `x` and heart weight as `y`. What is the estimate for $\beta$ for this data?

**(d)** Check your work in `R`. The following syntax can be used to fit a model without an intercept:

```{r, eval = FALSE}
lm(response ~ 0 + predictor, data = dataset)
```

Use this to fit a model to the `cat` data without an intercept. Output the coefficient of the fitted model. It should match your answer to **(c)**.

