---
title: "Untitled"
author: "Anh Nguyen"
date: "7/8/2020"
output: html_document
---

# Dianostic in R
Similating data

```{r}
# no sumption violation
sim_1 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x + rnorm(n = sample_size, mean = 0, sd = 1)
  data.frame(x, y)
}

sim_2 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x + rnorm(n = sample_size, mean = 0, sd = x)
  data.frame(x, y)
}

#not linear function
sim_3 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x ^ 2 + rnorm(n = sample_size, mean = 0, sd = 5)
  data.frame(x, y)
}
```

generate good data

```{r}
set.seed(42)
sim_data_1 = sim_1()
head(sim_data_1)
```

Scatterplot

```{r}
plot(y ~ x, data = sim_data_1, col = "grey", pch = 20,
     main = "Data from Model 1")
fit_1 = lm(y ~ x, data = sim_data_1)
abline(fit_1, col = "darkorange", lwd = 3)
```

```{r}
#Model 2
set.seed(42)
sim_data_2 = sim_2()
fit_2 = lm(y ~ x, data = sim_data_2)
plot(y ~ x, data = sim_data_2, col = "grey", pch = 20,
     main = "Data from Model 2")
abline(fit_2, col = "darkorange", lwd = 3)
```

```{r}
#model 3
set.seed(42)
sim_data_3 = sim_3()
fit_3 = lm(y ~ x, data = sim_data_3)
plot(y ~ x, data = sim_data_3, col = "grey", pch = 20,
     main = "Data from Model 3")
abline(fit_3, col = "darkorange", lwd = 3)
```


## Fitted vs residual plot

```{r}
plot(fitted(fit_1), resid(fit_1), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Data from Model 1")
abline(h = 0, col = "darkorange", lwd = 2)
```
```{r}
#model 2
plot(fitted(fit_2), resid(fit_2), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Data from Model 2")
abline(h = 0, col = "darkorange", lwd = 2)
```


```{r}
#Model 3
plot(fitted(fit_3), resid(fit_3), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Data from Model 3")
abline(h = 0, col = "darkorange", lwd = 2)
```

## breusch_pagan test

```{r}

library(lmtest)

bptest(fit_1)
```

## Histogram 


```{r}
par(mfrow = c(1, 3))
hist(resid(fit_1),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_1",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
hist(resid(fit_2),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_2",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
hist(resid(fit_3),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_3",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
```

## Q-Q Plot

### Function for QQ plot
```{r}
qq_plot = function(e) {

  n = length(e)
  normal_quantiles = qnorm(((1:n - 0.5) / n))
  # normal_quantiles = qnorm(((1:n) / (n + 1)))

  # plot theoretical verus observed quantiles
  plot(normal_quantiles, sort(e),
       xlab = c("Theoretical Quantiles"),
       ylab = c("Sample Quantiles"),
       col = "darkgrey")
  title("Normal Q-Q Plot")

  # calculate line through the first and third quartiles
  slope     = (quantile(e, 0.75) - quantile(e, 0.25)) / (qnorm(0.75) - qnorm(0.25))
  intercept = quantile(e, 0.25) - slope * qnorm(0.25)

  # add to existing plot
  abline(intercept, slope, lty = 2, lwd = 2, col = "dodgerblue")
}
```

### R function for QQ plot

```{r}
set.seed(420)
x = rnorm(100, mean = 0 , sd = 1)
par(mfrow = c(1, 2))
qqnorm(x, col = "darkgrey")
qqline(x, lty = 2, lwd = 2, col = "dodgerblue")
qq_plot(x)
```

Normal sample

```{r}
par(mfrow = c(1, 3))
set.seed(420)
qq_plot(rnorm(10))
qq_plot(rnorm(25))
qq_plot(rnorm(100))
```

t-dist
```{r}
par(mfrow = c(1, 3))
set.seed(420)
qq_plot(rt(10, df = 4))
qq_plot(rt(25, df = 4))
qq_plot(rt(100, df = 4))
```

Exponent distribution (skew)

```{r}
par(mfrow = c(1, 3))
set.seed(420)
qq_plot(rexp(10))
qq_plot(rexp(25))
qq_plot(rexp(100))
```

Q-Q plot for above data

fit 1
```{r}
qqnorm(resid(fit_1), main = "Normal Q-Q Plot, fit_1", col = "darkgrey")
qqline(resid(fit_1), col = "dodgerblue", lwd = 2)
```

fit 2

```{r}
qqnorm(resid(fit_2), main = "Normal Q-Q Plot, fit_3", col = "darkgrey")
qqline(resid(fit_3), col = "dodgerblue", lwd = 2)
```

```{r}
qqnorm(resid(fit_3), main = "Normal Q-Q Plot, fit_3", col = "darkgrey")
qqline(resid(fit_3), col = "dodgerblue", lwd = 2)
```

### Shapiro-wilk test

```{r}
set.seed(42)
shapiro.test(rnorm(25))
#gerenate data for non-normal distribution
shapiro.test(rexp(25))
```

Testing shapiro-wilk using residuals

```{r}
shapiro.test(resid(fit_1))
shapiro.test(resid(fit_2))
shapiro.test(resid(fit_3))
```



# Unsual Observation

```{r}
par(mfrow = c(1, 3))
set.seed(42)
ex_data  = data.frame(x = 1:10,
                      y = 10:1 + rnorm(n = 10))
ex_model = lm(y ~ x, data = ex_data)

# low leverage, large residual, small influence
point_1 = c(5.4, 11)
ex_data_1 = rbind(ex_data, point_1)
model_1 = lm(y ~ x, data = ex_data_1)
plot(y ~ x, data = ex_data_1, cex = 2, pch = 20, col = "grey",
     main = "Low Leverage, Large Residual, Small Influence")
points(x = point_1[1], y = point_1[2], pch = 1, cex = 4, col = "black", lwd = 2)
abline(ex_model, col = "dodgerblue", lwd = 2)
abline(model_1, lty = 2, col = "darkorange", lwd = 2)
legend("bottomleft", c("Original Data", "Added Point"),
       lty = c(1, 2), col = c("dodgerblue", "darkorange"))

# high leverage, small residual, small influence
point_2 = c(18, -5.7)
ex_data_2 = rbind(ex_data, point_2)
model_2 = lm(y ~ x, data = ex_data_2)
plot(y ~ x, data = ex_data_2, cex = 2, pch = 20, col = "grey",
     main = "High Leverage, Small Residual, Small Influence")
points(x = point_2[1], y = point_2[2], pch = 1, cex = 4, col = "black", lwd = 2)
abline(ex_model, col = "dodgerblue", lwd = 2)
abline(model_2, lty = 2, col = "darkorange", lwd = 2)
legend("bottomleft", c("Original Data", "Added Point"),
       lty = c(1, 2), col = c("dodgerblue", "darkorange"))

# high leverage, large residual, large influence
point_3 = c(14, 5.1)
ex_data_3 = rbind(ex_data, point_3)
model_3 = lm(y ~ x, data = ex_data_3)
plot(y ~ x, data = ex_data_3, cex = 2, pch = 20, col = "grey", ylim = c(-3, 12),
     main = "High Leverage, Large Residual, Large Influence")
points(x = point_3[1], y = point_3[2], pch = 1, cex = 4, col = "black", lwd = 2)
abline(ex_model, col = "dodgerblue", lwd = 2)
abline(model_3, lty = 2, col = "darkorange", lwd = 2)
legend("bottomleft", c("Original Data", "Added Point"),
       lty = c(1, 2), col = c("dodgerblue", "darkorange"))
```

## Leverage
```{r}
hatvalues(model_1)
hatvalues(model_2)
hatvalues(model_3)

```

### Find the high leverage
```{r}
hatvalues(model_1) > 2 * mean(hatvalues(model_1))
hatvalues(model_2) > 2 * mean(hatvalues(model_2))
hatvalues(model_3) > 2 * mean(hatvalues(model_3))
```

## Outliners
### Residual

```{r}
resid(model_1)
resid(model_2)
resid(model_3)
```

### Standardized residuals

```{r}
rstandard(model_1)
rstandard(model_2)
rstandard(model_3)

```

### Detect outliners

```{r}
rstandard(model_1)[abs(rstandard(model_1)) > 2]
rstandard(model_2)[abs(rstandard(model_2)) > 2]
rstandard(model_3)[abs(rstandard(model_3)) > 2]
```



## Cook's distance
```{r}
cooks.distance(model_1)[11] > 4 / length(cooks.distance(model_1))

cooks.distance(model_2)[11] > 4 /length(cooks.distance(model_2))

cooks.distance(model_3)[11] > 4 / length(cooks.distance(model_3))
```

# mtcars example

```{r}
mpg_hp_add = lm(mpg ~ hp + am, data = mtcars)
```

```{r}
plot(fitted(mpg_hp_add), resid(mpg_hp_add), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residual",
     main = "mtcars: Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
```

bp-test 

```{r}
bptest(mpg_hp_add)
```

Q-Q plot

```{r}
qqnorm(resid(mpg_hp_add), col = "darkgrey")
qqline(resid(mpg_hp_add), col = "dodgerblue", lwd = 2)
```

Shapiro.test
```{r}
shapiro.test(resid(mpg_hp_add))
```

Finding high leverage point

```{r}
sum(hatvalues(mpg_hp_add) > 2 * mean(hatvalues(mpg_hp_add)))
```

## Finding large residual point

```{r}
sum(abs(rstandard(mpg_hp_add)) > 2)
```

##Check influence point

```{r}
cd_mpg_hp_add = cooks.distance(mpg_hp_add)
sum(cd_mpg_hp_add > 4 / length(cd_mpg_hp_add))
```

```{r}
large_cd_mpg = cd_mpg_hp_add > 4 / length(cd_mpg_hp_add)
cd_mpg_hp_add[large_cd_mpg]
```

## Removing influence points

```{r}
mpg_hp_add_fix = lm(mpg ~ hp + am,
                    data = mtcars,
                    subset = cd_mpg_hp_add <= 4 / length(cd_mpg_hp_add))
coef(mpg_hp_add_fix)
```
## Plot
```{r}
par(mfrow = c(2, 2))
plot(mpg_hp_add)
```

# Transformation

Input data

```{r}
initech = read.csv("initech.csv")
```

Plotting data
```{r}

plot(salary ~ years, data = initech, col = "grey", pch = 20, cex = 1.5,
     main = "Salaries at Initech, By Seniority")
```

```{r}
initech_fit = lm(salary ~ years, data = initech)
summary(initech_fit)
```

```{r}
plot(salary ~ years, data = initech, col = "grey", pch = 20, cex = 1.5,
     main = "Salaries at Initech, By Seniority")
abline(initech_fit, col = "darkorange", lwd = 2)
```

```{r fig.height=5, fig.width=10}
par(mfrow = c(1, 2))

plot(fitted(initech_fit), resid(initech_fit), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(initech_fit), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(initech_fit), col = "dodgerblue", lwd = 2)
```

## Log transform - Response Transform

```{r}
initech_fit_log = lm(log(salary) ~ years, data = initech)
```

```{r}
plot(log(salary) ~ years, data = initech, col = "grey", pch = 20, cex = 1.5,
     main = "Salaries at Initech, By Seniority")
abline(initech_fit_log, col = "darkorange", lwd = 2)
```

Checking 2 assumption
```{r}
par(mfrow = c(1, 2))

plot(fitted(initech_fit_log), resid(initech_fit_log), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(initech_fit_log), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(initech_fit_log), col = "dodgerblue", lwd = 2)
```
### Transform back to exp(bo + b1x1)

Adding curve
```{r}
plot(salary ~ years, data = initech, col = "grey", pch = 20, cex = 1.5,
     main = "Salaries at Initech, By Seniority")
curve(exp(initech_fit_log$coef[1] + initech_fit_log$coef[2] * x), #beta0_hat + beta1_hat*x
      from = 0, to = 30, add = TRUE, col = "darkorange", lwd = 2)
```

### RMSE (checking which model is better)

```{r}
# Before transform
sqrt(mean((initech$salary - fitted(initech_fit)) ^ 2))
#after transform
sqrt(mean((initech$salary - exp(fitted(initech_fit_log))) ^ 2))
```

## Predictor Transform

### Quadratic model

Similating function
```{r}
sim_quad = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x ^ 2 + rnorm(n = sample_size, mean = 0, sd = 5)
  data.frame(x, y)
}
```

Simulating Data

```{r}
set.seed(314)
quad_data = sim_quad(sample_size = 200)
```

Fitted Model
```{r}
lin_fit = lm(y ~ x, data = quad_data)
summary(lin_fit)
```

```{r}
plot(y ~ x, data = quad_data, col = "grey", pch = 20, cex = 1.5,
     main = "Simulated Quadratic Data")
abline(lin_fit, col = "darkorange", lwd = 2)
```

$$
Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \epsilon_i
$$


Fitting model with x^2 term

```{r}
quad_fit = lm(y ~ x + I(x^2), data = quad_data)
summary(quad_fit)
```

Plotting using curve function
```{r}
plot(y ~ x, data = quad_data, col = "grey", pch = 20, cex = 1.5,
     main = "Simulated Quadratic Data")
curve(quad_fit$coef[1] + quad_fit$coef[2] * x + quad_fit$coef[3] * x ^ 2,
      from = -5, to = 30, add = TRUE, col = "darkorange", lwd = 2)
```

```{r, fig.height=5, fig.width=10}
par(mfrow = c(1, 2))

plot(fitted(quad_fit), resid(quad_fit), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(quad_fit), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(quad_fit), col = "dodgerblue", lwd = 2)
````
### Overfitting and Extrapolation

```{r}
sim_for_perf = function() {
  x = seq(0, 10)
  y = 3 + x - 4 * x ^ 2 + rnorm(n = 11, mean = 0, sd = 25)
  data.frame(x, y)
}
```

```{r}
set.seed(1234)
data_for_perf = sim_for_perf()
```

Fit_correct: Correct form
```{r}
fit_correct = lm(y ~ x + I(x ^ 2), data = data_for_perf)
fit_perfect = lm(y ~ x + I(x ^ 2) + I(x ^ 3) + I(x ^ 4) + I(x ^ 5) + I(x ^ 6) + 
                 I(x ^ 7) + I(x ^ 8) + I(x ^ 9) + I(x ^ 10), 
                 data = data_for_perf)
```

```{r, fig.height=6, fig.width=8}
x_plot = seq(-5, 15, by = 0.1) # Meke xvalue sequene
plot(y ~ x, data = data_for_perf, ylim = c(-450, 100), cex = 2, pch = 20)
lines(x_plot, predict(fit_correct, newdata = data.frame(x = x_plot)),
      col = "dodgerblue", lwd = 2, lty = 1)
lines(x_plot, predict(fit_perfect, newdata = data.frame(x = x_plot)),
      col = "darkorange", lwd = 2, lty = 2)
```
### Comparing Polynomial Models

```{r}
sim_higher = function(sample_size = 250) {
  x = runif(n = sample_size, min = -1, max = 1) * 2
  y = 3 + -6 * x ^ 2 + 1 * x ^ 4 + rnorm(n = sample_size, mean = 0, sd = 3)
  data.frame(x, y)
}
```

We are creating 3 y. 1 for quaratic term, 4th order term and 6th order terms
$$
Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \epsilon_i
$$

$$
Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \beta_3 x_i^3 + \beta_4 x_i^4 + \epsilon_i
$$

$$
Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \beta_3 x_i^3 + \beta_4 x_i^4 + \beta_5 x_i^5 + \beta_6 x_i^6 + \epsilon_i
$$

```{r}
set.seed(42)
data_higher = sim_higher()
```

```{r}
plot(y ~ x, data = data_higher, col = "grey", pch = 20, cex = 1.5,
     main = "Simulated Quartic Data")
```

Using Poly function instead of I function
```{r}
fit_2 = lm(y ~ poly(x, 2), data = data_higher)
fit_4 = lm(y ~ poly(x, 4), data = data_higher)
```

```{r}
plot(y ~ x, data = data_higher, col = "grey", pch = 20, cex = 1.5,
     main = "Simulated Quartic Data")
x_plot = seq(-5, 5, by = 0.05)
lines(x_plot, predict(fit_2, newdata = data.frame(x = x_plot)),
      col = "dodgerblue", lwd = 2, lty = 1)
lines(x_plot, predict(fit_4, newdata = data.frame(x = x_plot)),
      col = "darkorange", lwd = 2, lty = 2)
```

dianoge for quaratic Model
```{r, fig.height=5, fig.width=10}
par(mfrow = c(1, 2))

plot(fitted(fit_2), resid(fit_2), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(fit_2), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(fit_2), col = "dodgerblue", lwd = 2)
```

Dianose for 4th Model
```{r, fig.height=5, fig.width=10}
par(mfrow = c(1, 2))

plot(fitted(fit_4), resid(fit_4), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

qqnorm(resid(fit_4), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(fit_4), col = "dodgerblue", lwd = 2)
```
#### Testing model

```{r}
anova(fit_2, fit_4)
```
-> Prefer larger model

Testing 6th power model
```{r}
fit_6 = lm(y ~ poly(x, 6), data = data_higher)
```

```{r}
anova(fit_4, fit_6)
```

### Poly function

$$
Y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \beta_3 x_i^3 + \beta_4 x_i^4 + \epsilon_i
$$

```{r}
fit_4a = lm(y ~ poly(x, degree = 4), data = data_higher)
fit_4b = lm(y ~ poly(x, degree = 4, raw = TRUE), data = data_higher)
fit_4c = lm(y ~ x + I(x^2) + I(x^3) + I(x^4), data = data_higher)
```

```{r}
coef(fit_4a)
coef(fit_4b)
coef(fit_4c)
```

To remove all the name in coef
```{r}
unname(coef(fit_4a))
unname(coef(fit_4b))
unname(coef(fit_4c))
```

```{r}
all.equal(fitted(fit_4a),
          fitted(fit_4b))
```

```{r}
all.equal(resid(fit_4a),
          resid(fit_4b))
```

```{r}
summary(fit_4a)
```

```{r}
summary(fit_4c)
```

### Inhibit Function

```{r}
coef(lm(y ~ x + x ^ 2, data = quad_data))
coef(lm(y ~ x + I(x ^ 2), data = quad_data))
```

```{r}
coef(lm(y ~ x + x:x, data = quad_data))
coef(lm(y ~ x * x, data = quad_data))
coef(lm(y ~ x ^ 2, data = quad_data))
coef(lm(y ~ x + x ^ 2, data = quad_data))
```

```{r}
coef(lm(y ~ I(x + x), data = quad_data))
coef(lm(y ~ x + x, data = quad_data))
```

### How to consiser which paranomial data
```{r, echo = FALSE}
# read data frame from the web
autompg = read.table(
  "http://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",
  quote = "\"",
  comment.char = "",
  stringsAsFactors = FALSE)
# give the dataframe headers
colnames(autompg) = c("mpg", "cyl", "disp", "hp", "wt", "acc", "year", "origin", "name")
# remove missing data, which is stored as "?"
autompg = subset(autompg, autompg$hp != "?")
# remove the plymouth reliant, as it causes some issues
autompg = subset(autompg, autompg$name != "plymouth reliant")
# give the dataset row names, based on the engine, year and name
rownames(autompg) = paste(autompg$cyl, "cylinder", autompg$year, autompg$name)
# remove the variable for name
autompg = subset(autompg, select = c("mpg", "cyl", "disp", "hp", "wt", "acc", "year", "origin"))
# change horsepower from character to numeric
autompg$hp = as.numeric(autompg$hp)
# create a dummary variable for foreign vs domestic cars. domestic = 1.
autompg$domestic = as.numeric(autompg$origin == 1)
# remove 3 and 5 cylinder cars (which are very rare.)
autompg = autompg[autompg$cyl != 5,]
autompg = autompg[autompg$cyl != 3,]
# change cyl to a factor variable
autompg$cyl = as.factor(autompg$cyl)
```

```{r, fig.height=10, fig.width=10}
pairs(autompg)
```

Fit SLR

```{r, fig.height=5, fig.width=10}
mpg_hp = lm(mpg ~ hp, data = autompg)

par(mfrow = c(1, 2))

plot(mpg ~ hp, data = autompg, col = "dodgerblue", pch = 20, cex = 1.5)
abline(mpg_hp, col = "darkorange", lwd = 2)

plot(fitted(mpg_hp), resid(mpg_hp), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

```{r, fig.height=5, fig.width=10}
mpg_hp_log = lm(mpg ~ hp + I(hp ^ 2), data = autompg)

par(mfrow = c(1, 2))

plot(mpg ~ hp, data = autompg, col = "dodgerblue", pch = 20, cex = 1.5)
xplot = seq(min(autompg$hp), max(autompg$hp), by = 0.1)
lines(xplot, predict(mpg_hp_log, newdata = data.frame(hp = xplot)),
      col = "darkorange", lwd = 2, lty = 1)

plot(fitted(mpg_hp_log), resid(mpg_hp_log), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

```{r, fig.height=5, fig.width=10}
mpg_hp_log = lm(log(mpg) ~ hp + I(hp ^ 2), data = autompg)

par(mfrow = c(1, 2))

plot(log(mpg) ~ hp, data = autompg, col = "dodgerblue", pch = 20, cex = 1.5)
xplot = seq(min(autompg$hp), max(autompg$hp), by = 0.1)
lines(xplot, predict(mpg_hp_log, newdata = data.frame(hp = xplot)),
      col = "darkorange", lwd = 2, lty = 1)

plot(fitted(mpg_hp_log), resid(mpg_hp_log), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

```{r, fig.height=5, fig.width=10}
mpg_hp_loglog = lm(log(mpg) ~ log(hp), data = autompg)

par(mfrow = c(1, 2))
plot(log(mpg) ~ log(hp), data = autompg, col = "dodgerblue", pch = 20, cex = 1.5)
abline(mpg_hp_loglog, col = "darkorange", lwd = 2)

plot(fitted(mpg_hp_loglog), resid(mpg_hp_loglog), col = "dodgerblue",
     pch = 20, cex = 1.5, xlab = "Fitted", ylab = "Residuals")
abline(h = 0, lty = 2, col = "darkorange", lwd = 2)
```

```{r}
big_model = lm(mpg ~ disp * hp * domestic, data = autompg)
```

```{r}
qqnorm(resid(big_model), col = "darkgrey")
qqline(resid(big_model), col = "dodgerblue", lwd = 2)
```

```{r}
bigger_model = lm(log(mpg) ~ disp * hp * domestic + 
               I(disp ^ 2) + I(hp ^ 2), data = autompg)
summary(bigger_model)
```

```{r}
qqnorm(resid(bigger_model), col = "darkgrey")
qqline(resid(bigger_model), col = "dodgerblue", lwd = 2)
```