---
title: "School's document"
author: "Anh Nguyen"
date: "6/03/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Variance of Beta Hat

\[
\hat{\beta}_1 \sim N\left(  \beta_1, \frac{\sigma^2}{S_{xx}} \right)
\]

and

\[
\hat{\beta}_0 \sim N\left(  \beta_0, \sigma^2 \left(\frac{1}{n} + \frac{\bar{x}^2}{S_{xx}}\right) \right).
\]

First we need to decide ahead of time what our $x$ values will be for this simulation, since the $x$ values in SLR are also considered known quantities. The choice of $x$ values is arbitrary. Here we also set a seed for randomization, and calculate $S_{xx}$ which we will need going forward.

```{r}
set.seed(42)
sample_size = 100 # this is n
x = seq(-1, 1, length = sample_size)
Sxx = sum((x - mean(x)) ^ 2)
```

We also fix our parameter values.

```{r}
beta_0 = 3
beta_1 = 6
sigma  = 2
```

With this information, we know the sampling distributions should be:

```{r}
#True variance of beta_1_hat 
(var_beta_1_hat = sigma ^ 2 / Sxx)
(var_beta_0_hat = sigma ^ 2 * (1 / sample_size + mean(x) ^ 2 / Sxx))
```

\[
\hat{\beta}_1 \sim N(  `r beta_1`, `r var_beta_1_hat`)
\]

and

\[
\hat{\beta}_0 \sim N(  `r beta_0`, `r var_beta_0_hat`).
\]

That is,

\[
\begin{aligned}
\text{E}[\hat{\beta}_1] &= `r beta_1` \\
\text{Var}[\hat{\beta}_1] &= `r var_beta_1_hat`
\end{aligned}
\]

and

\[
\begin{aligned}
\text{E}[\hat{\beta}_0] &= `r beta_0` \\
\text{Var}[\hat{\beta}_0] &= `r var_beta_0_hat`.
\end{aligned}
\]

# Variance of B1 hat

Now, simulate data from this model 10000 times

```{r}
#preallocate the vector
num_samples = 10000
beta_0_hats = rep(0, num_samples)
beta_1_hats = rep(0, num_samples)

for (i in 1:num_samples) {
  eps = rnorm(sample_size, mean = 0, sd = sigma) #noise using sigma
  y   = beta_0 + beta_1 * x + eps #generate y value for this dataset
  
  sim_model = lm(y ~ x) #fit regression
  
  beta_0_hats[i] = coef(sim_model)[1] #extract beta0 & beta 1
  beta_1_hats[i] = coef(sim_model)[2]
}
```


Each time we simulated the data, we obtained values of the estimated coefficiets. The variables `beta_0_hats` and `beta_1_hats` now store 10,000 simulated values of $\hat{\beta}_0$ and $\hat{\beta}_1$ respectively.

We first verify the distribution of $\hat{\beta}_1$.

```{r}
mean(beta_1_hats) # empirical mean
beta_1            # true mean
var(beta_1_hats)  # empirical variance
var_beta_1_hat    # true variance
```

We see that the empirical and true means and variances are *very* similar. We also verify that the empirical distribution is normal. To do so, we plot a histogram of the `beta_1_hats`, and add the curve for the true distribution of $\hat{\beta}_1$. We use `prob = TRUE` to put the histogram on the same scale as the normal curve.

# Verifying distribution NORMAL

## Histogram

```{r}
hist(beta_1_hats, prob = TRUE, breaks = 25, 
     xlab = expression(hat(beta)[1]), 
     main = "",
     border = "dodgerblue")

# Adding curve for normal distribution
curve(dnorm(x, mean = beta_1, sd = sqrt
            (var_beta_1_hat)),
      col = "darkorange", add = TRUE, lwd = 3)
```


## Check Probability

```{r}
mean(beta_1_hats <6.2)
pnorm (6.2, mean = beta_1, sd = sqrt(var_beta_1_hat))
```

## Plot

```{r}
par(mar = c(5, 5, 1, 1)) # adjusted plot margins, otherwise the "hat" does not display
plot(cumsum(beta_1_hats) / (1:length(beta_1_hats)), type = "l", ylim = c(5.95, 6.05),
     xlab = "Number of Simulations",
     ylab = expression("Empirical Mean of " ~ hat(beta)[1]),
     col  = "dodgerblue")
abline(h = 6, col = "darkorange", lwd = 2)

par(mar = c(5, 5, 1, 1)) # adjusted plot margins, otherwise the "hat" does not display
plot(cumsum(beta_0_hats) / (1:length(beta_0_hats)), type = "l", ylim = c(2.95, 3.05),
     xlab = "Number of Simulations",
     ylab = expression("Empirical Mean of " ~ hat(beta)[0]),
     col  = "dodgerblue")
abline(h = 3, col = "darkorange", lwd = 2)

```

#Testing in R


```{r}
# fit model
stop_dist_model = lm(dist ~ speed, data = cars)

#Summary fitted model
summary(stop_dist_model)
```

```{r}
names(summary(stop_dist_model))

```

The `names()` function tells us what information is available, and then we use the `$` operator and `coefficients` to extract the information we are interested in. Two values here should be immediately familiar.

```{r}
summary(stop_dist_model)$coefficients
```

- Estimate : beta^1, beta^0 (Estimated value of beta 1 and beta 0)
- Std Error: SD of beta1 and beta0
- T value
- P- valude

\[
\hat{\beta}_0 = `r summary(stop_dist_model)$coefficients[1,1]`
\]

and

\[
\hat{\beta}_1 = `r summary(stop_dist_model)$coefficients[2,1]`
\]
## Plot data
```{r}
plot(dist ~ speed, data = cars,
     xlab = "Speed (in Miles per hours",
     ylab = "Stopping Distance (in feet)",
     main = "Stopping Distance vs Speed",
     pch = 20,
     cex = 2,
     col = "grey")

abline(stop_dist_model, lwd = 5, col = "orange")
```



## Value of B^1


```{r}
#Getting full sumary of B^1
summary(stop_dist_model)$coefficients[2,]
```

Again, the first value, `Estimate` is

\[
\hat{\beta}_1 = `r summary(stop_dist_model)$coefficients[2,1]`.
\]

The second value, `Std. Error`, is the standard error of $\hat{\beta}_1$,

\[
\text{SE}[\hat{\beta}_1] = \frac{s_e}{\sqrt{S_{xx}}} = `r summary(stop_dist_model)$coefficients[2,2]`.
\]

The third value, `t value`, is the value of the test statistic for testing $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$,

\[
t = \frac{\hat{\beta}_1-0}{\text{SE}[\hat{\beta}_1]} = \frac{\hat{\beta}_1-0}{s_e / \sqrt{S_{xx}}} = `r summary(stop_dist_model)$coefficients[2,3]`.
\]

Lastly, `Pr(>|t|)`, gives us the p-value of that test.

\[
\text{p-value} = `r summary(stop_dist_model)$coefficients[2,4]`
\]

## Value of B^0

```{r}
summary(stop_dist_model)$coefficients[1,]
```


## Storing B^1, B^0 values

```{r}
stop_dist_model_test_info = summary(stop_dist_model)$coefficients

beta_0_hat      = stop_dist_model_test_info[1, 1] # Estimate
beta_0_hat_se   = stop_dist_model_test_info[1, 2] # Std. Error
beta_0_hat_t    = stop_dist_model_test_info[1, 3] # t value
#beta_0_hat_t = sumary(stop_dist_model)$coefficient["(Intercept)","t_value"]

beta_0_hat_pval = stop_dist_model_test_info[1, 4] # Pr(>|t|)

beta_1_hat      = stop_dist_model_test_info[2, 1] # Estimate
beta_1_hat_se   = stop_dist_model_test_info[2, 2] # Std. Error
beta_1_hat_t    = stop_dist_model_test_info[2, 3] # t value
#beta_1_hat_t = summary(stop_dist_model)$coefficients["speed","t_value"]
beta_1_hat_pval = stop_dist_model_test_info[2, 4] # Pr(>|t|)
```

##test statistic

```{r}
#b^0

summary(stop_dist_model)$coefficients["(Intercept)", "t-value"]
```


##P-value

```{r}
2 * pt(-abs(beta_1_hat_t), df = nrow(cars) -2)

2 * pt(abs(beta_1_hat_t), df = nrow(cars) -2, lower.tail = FALSE)
```

# Stimulate SLR

Function to simulate
```{r}
sim_slt = function(x, beta_0 = 10, beta_1 = 5, sigma = 1){
  n = length(x)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  data.frame(predictor = x, respond = y)
}
```

```{r}
#simulate SLR
set.seed(1)
x = seq(1,20, length.out = 21)

sim_data = sim_slt(x = x, beta_0 = 2, beta_1 = 0, sigma = 1)
```

```{r}
#fit SLR

sim_fit = lm(respond ~ predictor, data = sim_data)
summary(sim_fit)$coefficients["predictor","Pr(>|t|)"]
```



# Standard Error

```{r}
summary(stop_dist_model)$coefficients[,2]

#Beta 1 hat error

summary(stop_dist_model)$coefficients[2,2]

#beta 0 hat error

summary(stop_dist_model)$coefficients[1,2]
```


Verifying Standard Error
```{r}
Sxx = sum((cars$speed - mean(cars$speed))^2)
s_e = summary(stop_dist_model)$sigma
s_e/sqrt(Sxx)
```

```{r}
s_e/Sxx == beta_1_hat_se
```

#Critical Value - t - qt()

```{r}

#calculate critical value for 2 sides 99% CI

(1 - 0.99)/2

crit = qt(0.995, df = nrow(cars) - 2)
```

## Difference qt() vs pt()

pt is probability for t

```{r}
pt(qt(0.995, df = nrow(cars) - 2), df = nrow(cars) - 2)
```

getting alpha/2

```{r}
1 - pt(qt(0.995, df = nrow(cars) - 2), df = nrow(cars) - 2)
```


# Confident Interval - confint()

Computes confidence intervals in FITTED model

```{r}
confint(stop_dist_model, level = 0.99)
```

```{r}
# Navigate way to get a specific bounce
confint(stop_dist_model, level = 0.99)[2,2]
```


Using parm = to extract a particular parameters
```{r}
# confidence interval for particilar regression parameters
confint(stop_dist_model, parm = "(Intercept)" ,level = 0.99) #estimate for Beta 0

confint(stop_dist_model, parm = "speed" ,level = 0.99) #estimate for Beta 1
```

Verifying against the normal calculation


```{r}
#information about fitted model

summary(stop_dist_model)
```

```{r}

```

## Getting lower bounce (beta hat 1)
EST - CRIT * SE

```{r}
lower_bounce_beta1hat = beta_1_hat - crit * beta_1_hat_se
```

## Getting higher bounce (beta hat 1)
EST + CRIT * SE

```{r}
higher_bounce_beta1hat = beta_1_hat - crit * beta_1_hat_se
```

# ConInv for Mean respond

```{r}
#new data
new_speed = data.frame(speed = c(5,21))
```

```{r}
#point estimate
predict(stop_dist_model, newdata = new_speed)
```


```{r}
# Interval for mean of y at an x of 21
predict(stop_dist_model, newdata = new_speed,
        interval = c("confidence"), level = 0.99
        )
```

fit value is point estimate.

# Predicion Interval for new Observation - Predict()

```{r}
# Interval we hope to capture new observation at x of 21
predict(stop_dist_model, newdata = new_speed,
        interval = c("prediction"), level = 0.99
        )
```






